---
title: "Bird Count Analysis"
author: "Colin Scotland"
date: "26/03/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(here)
library(data.table)


here::here()
```

```{r}
source("data_cleaning_scripts/birds_cleaning.R")
```

### Analysis

For each question where we ask for bird names, give the bird’s common name, scientific name and species abbreviation.

#### 1. Which bird had the most individual sightings? 

```{r}
# Group the results by common_name;
  
birds_clean %>% 
  
  group_by(common_name, scientific_name, species_abbreviation) %>% 
  
  summarise(sight_count = n()) %>% 
  
  arrange(desc(sight_count))
  

  

  
  
```

The bird with the most sightings was the *Wandering albatross sensu lato*, scientific 
name *Diomedea antipodensis / exulans* with species abbreviation *DIOANTEXU*.  This
particular species had 11247 sightings.

#### 2. Which bird had the highest total count?

```{r}
birds_clean %>% 
  
  group_by(common_name, scientific_name, species_abbreviation) %>% 
  
  summarise(total_count = sum(count)) %>% 
  
  arrange(desc(total_count))


```

The bird with the highest total count is the *Short-tailed shearwater, Puffinus tenuirostris*,
or *PUFTEN* with a total count of 982553.

###### NOTE:
It was noted within the cleaning process that some species had very high counts of 99,999.

Further research, specifically on *Puffinus tenuirostris* showed that they did indeed exist 
in very large colonies, possibly exceeding 99,999 birds in multiple locations 
(see: *http://www.marineornithology.org/PDF/24/24_1.pdf*).  This report suggests 
that some protected areas off Tasmania hold well over 1 million nest burrows, so the result
above could actually be significantly lower than the true figure.

With this information it seemed prudent to keep the 99,999 values unchanged, with an 
assumption that this was the upper limit of the count that could be recorded at any one time.

--------------------------------------------------------------------------------


#### 3. Which bird had the highest total count above a latitude of -30?

```{r}
birds_clean %>% 
  
  filter(lat > -30) %>% 
  
  group_by(common_name, 
           scientific_name, 
           species_abbreviation) %>% 
  
  summarise(total_count = sum(count)) %>% 
  
  arrange(desc(total_count))
```

The bird with the highest total count above latitude -30 was the *Wedge-tailed shearwater*, 
*Puffinus pacificus* or PUFPAC, with a total count of 855.


#### 4. How many different types of birds were only ever seen in groups of 1?

i.e. how many species of birds were **only** ever seen alone.  One way to determine this would be
to generate a list of all the birds with a count of 1, then generate a list with a count > 1
and to `anti_join` the two data sets so that any matches in species were removed.  

```{r}
solo_birds <- birds_clean %>% 
  
  filter(count == 1) %>% 
  
  group_by(common_name) %>% 
  
  summarise()
  
```

```{r}
flock_birds <- birds_clean %>% 
  
  filter(count > 1) %>% 
  
  group_by(common_name) %>% 
  
  summarise()
```

The two lists have now been generated; they should be joined in such a way that 
only those birds from `solo_birds` that are **not** in `flock_birds` are returned.

An `anti_join` will keep values in 'x' that do not have a match in 'y';

```{r}
anti_join(solo_birds, flock_birds, by = "common_name")
```


This suggests that there were 26 bird species that were **only** ever sighted as 
solitary birds, *but* there  are still some capitalised suffixes on some common names 
here which could be missed during the anti-join, so re-run the regex again; 

```{r}
solo_birds <- solo_birds %>% 
  
  mutate(common_name = str_remove(common_name, "\\s[A-Z]*[0-9]*$"))

flock_birds <- flock_birds %>% 
  
  mutate(common_name = str_remove(common_name, "\\s[A-Z]*[0-9]*$"))
```

...and redo the anti-join;

```{r}
anti_join(solo_birds, flock_birds, by = "common_name")
```

This suggests there are **23 species** that were **only** ever seen in solitude.


#### 5. How many penguins were seen? (Hint: there are many types of penguin)


```{r}
pattern <- " *[Pp]+[Ee]+[Nn]+[Gg]+[Uu]+[Ii]+[Nn] *"


penguins_by_species <- birds_clean %>% 
  
  filter(common_name %like% pattern) %>% 
  
  group_by(common_name) %>% 
  
  summarise(total_penguins_by_species = sum(count))

penguins_by_species

penguins_by_species %>% 
  
  summarise(sum(total_penguins_by_species))
  
  

```


In total **158 penguins** were recorded.
