---
title: "Halloween Sweets Analysis"
author: "Colin Scotland"
date: "29/03/2022"
output: html_document
---

The cleaning script for this analysis is `halloween_candy_cleaning_script.R`.  This
cleaning script can be found in the *data_cleaning_script* folder and runs the following;

1. reads in three .csv files containing "trick or treat" data for 2015, 2016 and 2017
2. removes unnecessary columns that are not needed for any analysis
3. corrects the time format for 2017 and reduces all three data sets to show only the year
4. changes the order of certain columns to set up for joining
5. renames several columns to make the join easier
6. performs a `full_join` to bring all the data into one data set while keeping all data
7. saves a clean data set to a .csv file in the *clean_data* folder
8. cleans up the environment

In summary, the file structure is;

* Top level --> .proj file
* data_cleaning_scripts --> .R script file
* clean_data --> clean .csv file
* documentation_and_analysis --> .Rmd and .html output file
* raw_data --> raw, uncleaned .csv/.xls files

The questions this analysis sets out to answer are;

1. What is the total number of candy ratings given across the three years. 
2. What was the average age of people who are going out trick or treating?
3. What was the average age of people who are not going trick or treating?
4. For each of joy, despair and meh, which candy bar revived the most of these ratings?
5. How many people rated Starburst as despair?
6. What was the most popular candy bar by this rating system for each gender in the dataset ?
7. What was the most popular candy bar in each year?
8. What was the most popular candy bar by this rating for people in US, Canada, UK, and all other countries?

A function `gender_joy()` was written in an attempt to streamline the code for question 6, 
and although the function "works", it is still incomplete.*

The results from the function are slightly different to the original analysis and the arguments
need to be set up correctly in terms of uppercase/lowercase.  This is still work in progress at
the time of submission to Github.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set up libraries;
```{r}
library(tidyverse)
library(here)
library(assertr)
library(testthat)
```

Set 'home' directory;
```{r}
here::here()
```

Source cleaning script;

```{r}
source(here("data_cleaning_scripts/halloween_candy_cleaning_script.R"))
```

## 1. What is the total number of candy ratings given across three years.

### (Number of candy ratings, not the number of raters.  Don't count missing values).

The first thing to do is to pivot the data into a longer form to allow easier analysis;

```{r}
# Add an ID column to keep track after pivoting;
candy_all <- candy_all %>% 
  mutate(id = 1:nrow(candy_all), .before = "year")


candy_long <- candy_all %>% 
  pivot_longer(cols = c(-id, -year, -age, -gender, -country, -going_out))

```

Then assign some values to *JOY*, *MEH* and *DESPAIR* to allow them to be counted;


```{r}
candy_long_ratings <- candy_long %>% 
  
  mutate(num_ratings = if_else(value == "(?i)JOY" |
                               value == "(?i)MEH" |
                               value == "(?i)DESPAIR", as.integer(1), as.integer(0)), .after = "value") %>% 
  
  summarise(total_ratings = sum(!is.na(num_ratings)))

candy_long_ratings
  
```

#### 1. There are **603883** ratings of sweets across the three years.  

These can also be broken down further to show how many of each type there were;

```{r}
candy_long %>% 
  
  select(value) %>% 
  
  distinct(value)
  
```

So there are only "JOY", "DESPAIR", "MEH" and NAs to deal with.

```{r}
candy_long_q1 <- candy_long %>% 
  
 # select(name, value) %>% 
  
  mutate(joy = (if_else(value == "JOY", 1, 0)), 
         despair = (if_else(value == "DESPAIR", 1, 0)), 
         meh = (if_else(value == "MEH", 1, 0)), 
         .after = "value") 

candy_long_q1 %>% 
  
  summarise(sum(joy, na.rm = TRUE), sum(despair, na.rm = TRUE), sum(meh, na.rm = TRUE))


```

Double check the count: 267672 + 260162 + 76049 = **603883**, which matches the original answer.


## 2.  What was the average age of people who are going out trick or treating?

As a start it would be good to have the age column formatted as integers;
```{r}
candy_all <- candy_all %>% 
  
  mutate(age = round(as.integer(age), digits = 1)) 
  

```
Note: the "junk" age entries have been coerced into NAs during the `as.integer` change, leaving
only numbers and NAs.  This is *a good thing* in this case.

In theory, a parent could take a baby trick or treating, and have answered the survey 
from their baby's perspective, so ages > 0 should be accepted.  At the other end of the 
spectrum, the oldest person in the world is reported to be 119 years old, so anyone aged 
120 or over is lying. 

```{r}
candy_all <- candy_all %>% 
  filter(age > 0 & age < 120)

# and check there are no surprises;


candy_all %>% 
  filter(!is.na(age)) %>% 
  verify(age > 0 & age < 120) 
```

Now, discounting NAs we want to know what the average age is of people going out.

```{r}
candy_all %>% 
  
# what are the responses - are they only 'yes'/'no', or are there some oddities or colloquialisms in there?
  
  distinct(going_out)

# There are only 'Yes', 'No' and 'NA'.

candy_all %>% 
  
  filter(going_out == "Yes") %>% 
  summarise(avg_age_out = mean(age, na.rm = TRUE))
```

#### 2. So the average age of people going out trick or treating is just over **35 years**.

## 3. What was the average age of people **not** going trick or treating?

Similarly to the last question;

```{r}
candy_all %>% 
  
  filter(going_out == "No") %>% 
  
  summarise(avg_age_out = mean(age, na.rm = TRUE))
```

#### 3. The average age of people going out trick or treating is just over **39 years**.

## 4. For each joy, despair and meh, which candy bar received the most of these ratings?

Basically, sum all the "JOY", "MEH" and "DESPAIR" counts for each candy type and return the top scorer;

```{r}

candy_rating_rates <- candy_long_q1 %>% 
  
  group_by(name, joy, meh, despair) %>% 
  
  summarise(all_the_joy = sum(as.integer(joy, na.rm = TRUE)), 
            all_the_meh = sum(as.integer(meh, na.rm = TRUE)), 
            all_the_despair = sum(as.integer(despair, na.rm = TRUE))
            )

candy_rating_rates %>% 
  
  select(name, all_the_joy) %>% 
  
  arrange(desc(all_the_joy))
  
candy_rating_rates %>% 
  
  select(name, all_the_meh) %>% 
  
  arrange(desc(all_the_meh))

candy_rating_rates %>% 
  
  select(name, all_the_despair) %>% 
  
  arrange(desc(all_the_despair))
 
  
```

#### 4. The candy bar with the most ratings of each kind were;


**JOY**: Any reese's peanut butter cups     7369 votes

**MEH**: Lollipops                          1570 votes

**DESPAIR**: Gum from baseball cards        7341 votes


## 5. How many people rated Starburst as despair?

Filter on "starburst" and count the "DESPAIR";

```{r}
candy_rating_rates %>% 
  
select(name, all_the_despair) %>% 
  
filter(name == "starburst" & all_the_despair > 0)
```

#### 5. There were **1990 votes of 'despair'** for Starburst.

## 6. What was the most popular candy bar by this rating system for each gender in the data set?

First, what entries do we have for gender;

```{r}
candy_long %>% 
  
  distinct(gender)
  
```
So we have to deal with 'NA', Male', 'Female', 'Other', and 'I'd rather not say'.

At least the formatting is consistent :)

```{r}
male_favourite <- candy_long_q1 %>% 
  
  select(gender, name, joy) %>% 
  
  filter(gender == "Male" & joy == 1)  %>% 
  
  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_male = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_male))

female_favourite <- candy_long_q1 %>% 
    
  select(gender, name, joy) %>% 
  
  filter(gender == "Female" & joy == 1) %>% 
  
  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_female = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_female))
  
other_favourite <- candy_long_q1 %>%

  select(gender, name, joy) %>% 
  
  filter(gender == "Other" & joy == 1) %>% 
  
  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_other = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_other))
  
nosay_favourite <- candy_long_q1 %>% 
  
  select(gender, name, joy) %>% 
  
  filter(gender == "I'd rather not say" & joy == 1) %>% 
  
  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_nosay = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_nosay))
  
na_favourite <- candy_long_q1 %>% 
  
  select(gender, name, joy) %>% 
  
  filter(is.na(gender) & joy == 1) %>% 

  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_na = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_na))
```

```{r}

slice_max(male_favourite, "total_joy_male", n = 1)

slice_max(female_favourite, order_by = "total_joy_female", n = 1)

slice_max(other_favourite, order_by = "total_joy_other", n = 1)

slice_max(nosay_favourite, order_by = "total_joy_nosay", n = 1)

slice_max(na_favourite, order_by = "total_joy_na", n = 1)



```

#### 6(i). So the most popular sweets by gender are;

**Male:**                  reese's peanut butter cups	            1545 votes

**Female:   **             reese's peanut butter cups	            837 votes

**Other: **                kit kat	                              34 votes

**Prefer not to say:**     reese's peanut butter cups	            69 votes

**NA: **                   reese's peanut butter cups	            4890 votes	    



###### *Now that I've got some results I've just realised that I haven't used the system *
###### *defined by the task of setting despair = -1, meh = 0 and joy = +1.  So I'll redo the *
###### *question properly now;* 

```{r}
candy_long_q6 <- candy_long %>% 
  
  # set up the scores for ratings;
  
  mutate(value_numeric = case_when(value == "JOY" ~ as.numeric(1),
                                   value == "MEH" ~ as.numeric(0),
                                   value == "DESPAIR" ~ as.numeric(-1)))

```

The most popular sweets will be the ones with the highest scores after adding up all
the joy (+1) and despair (-1).


```{r}
q6_male <- candy_long_q6 %>% 
  
  filter(gender == "Male")  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_male = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(total_joy_male))




q6_female <- candy_long_q6 %>% 
  
  filter(gender == "Female")  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_female = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(total_joy_female))




q6_other <- candy_long_q6 %>% 
  
  filter(gender == "Other")  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_other = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(total_joy_other))




q6_nosay <- candy_long_q6 %>% 
  
  filter(gender == "I'd rather not say")  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_nosay = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(total_joy_nosay)) 




q6_na <- candy_long_q6 %>% 
  
  filter(is.na(gender))  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_na = sum(value_numeric, na.rm = TRUE))  %>% 
  
  arrange(desc(total_joy_na))
```

That's the results grouped, so now to find the top scorers;

```{r}
slice_max(male_favourite, order_by = "q6_male", n = 1)

slice_max(female_favourite, order_by = "q6_female", n = 1)

slice_max(other_favourite, order_by = "q6_other", n = 1)

slice_max(nosay_favourite, order_by = "q6_nosay", n = 1)

slice_max(na_favourite, order_by = "q6_na", n = 1)
```

#### 6(ii). So the most popular sweets by gender using the *proper* system are;

**Male:**                  reese's peanut butter cups	            1545 votes

**Female:   **             reese's peanut butter cups	            837 votes

**Other: **                kit kat	                              34 votes

**Prefer not to say:**     reese's peanut butter cups	            69 votes

**NA: **                   reese's peanut butter cups	            4890 votes	 

...which is exactly the same as the result from *not* using the defined "system", 
so at least they're consistent!

--------------------------------------------------------------------------------

*It felt like there was a lot of repetition in writing the answer to this question,*
*so I tried writing a function that would do the work for me.*

*The function isn't quite working yet (input formats still to be sorted) and wasn't*
*100% functional when this project had to be submitted to Github...*


```{r}
#attempt to streamline the above function; couldn't get a return and ran out of time to figure it out.

# gender_joy <- function(choose_gender) {
  
#   if(!is.character(choose_gender)) { # control non-character entries
#    stop(call. = FALSE, "Error in input; select from 'male', 'female', 'other', 'prefer not to say' or 'na'!")  
#   }
 
#   if(choose_gender != "male" & choose_gender != "Male" & choose_gender != "MALE" &
#      choose_gender != "female" & choose_gender != "Female" & choose_gender != "FEMALE" &
#      choose_gender != "other" & choose_gender != "Other" & choose_gender != "OTHER" &
#      choose_gender != "na" & choose_gender != "Na" & choose_gender != "NA" &
#      choose_gender != "prefer not to say" & 
#      choose_gender != "Prefer not to say" & 
#      choose_gender != "PREFER NOT TO SAY")     {
#     stop(call. = FALSE, "Error in input; select from 'Male', 'Female', 'Other', 'Prefer not to say' or 'Na'!")
   
#    } else {

#  if(choose_gender != "na" & choose_gender != "Na" & choose_gender != "NA") {
    
#  q6_joy <- candy_long_q6 %>% 
  
#  select(gender, name, value_numeric) %>% 
    
#  filter(gender == str_to_title(choose_gender)) %>%  
  # str_to_title doesn't work for "Prefer not to say" - work in progress...
  
#  group_by(name) %>% 
    
#  summarise(total_joy = sum(value_numeric, na.rm = TRUE)) %>% 

#  arrange(desc(total_joy)) 

#  return(slice_max(q6_joy, order_by = "total_joy", n = 1))
  
#  } else {
    
#  q6_joy <- candy_long_q6 %>% 
      
#  filter(is.na(gender)) %>%
      
#  summarise(total_joy = sum(value_numeric, na.rm = TRUE)) %>% 

#  arrange(desc(total_joy)) 

#  return(slice_max(q6_joy, order_by = "total_joy", n = 1))
#        }
#    }
# }
```


```{r}
#test_that("incorrect input returns an error", {
#  expect_error(gender_joy("1999AD"))
#  expect_error(gender_joy(1234))
#  expect_error(gender_joy(TRUE))
#  expect_error(gender_joy("[a-zA-Z]"))
#})
```


## 7. What was the most popular candy bar in each year?

```{r}

candy_long_2015 <- candy_long_q6 %>% 
  
  filter(year == "2015") %>%  
  
  group_by(name) %>% 
  
  summarise(top_2015 = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_2015))

slice_max(candy_long_2015, order_by = "value_numeric", n = 1)
  
  
```

```{r}

candy_long_2016 <- candy_long_q6 %>% 
  
  filter(year == "2016") %>%  
  
  group_by(name) %>% 
  
  summarise(top_2016 = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_2016))

slice_max(candy_long_2016, order_by = "value_numeric", n = 1)

```

```{r}

candy_long_2017 <- candy_long_q6 %>% 
  
  filter(year == "2017") %>%  
  
  group_by(name) %>% 
  
  summarise(top_2017 = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_2017))

slice_max(candy_long_2017, order_by = "value_numeric", n = 1)
```

#### 7. So the most popular sweets by year are;

**2015:**   reese's peanut butter cups	   4375 votes

**2016:**   kit kat              	          920 votes

**2017:**   reese's peanut butter cups     1403 votes


## 8. What was the most popular candy bar by this rating for people in US, Canada,
UK and all other countries?

First thing to do is to clean up the 'country' column;

```{r}
#countries <- candy_long %>% 
  
#  distinct(country) %>% 
#  arrange(country)

# Lowercase country names for simpler regex;

countries <- candy_long_q6 %>% 
  mutate(country = str_to_lower(country))


  


# UK cleanup ----

countries <- countries %>% 
  mutate(country = str_replace_all(country, pattern = "[Ee]n[dg]land|[Ss]cotland|(?i)u[:punct:]*k[:punct:]*|(?i)united king*dom", "Great Britain")) 

# Canada cleanup ----
countries <- countries %>% 
  mutate(country = str_replace_all(country, pattern = "(?i)\\s*cana*d*a*e*[:symbol:]*|(?i)cascadia\\s*\\s*|(?i)Soviet Canuckistan", "Canada"))


# USA cleanup ----


countries <- countries %>% 
  
  mutate(country = str_remove_all(country, pattern = "(?i)\\sof\\samerica")) 

countries <- countries %>%                                     
  mutate(country = str_remove_all(country, pattern = "^USA[A-Za-z0-9\\s\\(\\)[:punct:][:symbol:]]{1,100}"))  

countries <- countries %>%                                    
  mutate(country = str_replace_all(country, pattern = "^[ahem]{0,4}[:punct:]{0,4}a*m+[eu]+r+[cir]+[aci]+[ak]*a*", "USA"))                             
                                    
countries <- countries %>%   
  mutate(country = str_replace_all(country, pattern = 
                                   "u+n+i+t+e+d+\\s+s+t+a+t+e+s+|^u+\\s*s{1,2}+\\s*a*",
                                   "USA")) 

countries <- countries %>%   
  mutate(country = str_remove_all(country, pattern = "^b+e+s+t+\\s+o+n+e+\\s+[:punct:]+\\s+"))


countries <- countries %>% 
  mutate(country = str_replace_all(country, pattern = "(?i)^u+s+a+[:punct:]*\\s*u*s*a*[:punct:]*\\s*u*s*a*[:punct:]*", "USA"))

countries <- countries %>% 
  mutate(country = str_replace_all(country, pattern = "(?i)^u+\\.*s+\\.*a*\\.*\\s*u*s*a*", "USA"))

countries <- countries %>% 
  mutate(country = str_replace_all(country, pattern = "(?i)i+\\s+p+r+e+t+e+n+d+\\s+[a-zA-Z\\s]{1,99}", "USA"))

countries <- countries %>% 
  mutate(country = str_replace_all(country, pattern = "(?i)^usa[A-Za-z0-9\\s\\,\\.\\'\\)]{1,80}", "USA"))

countries <- countries %>% 
  mutate(country = str_replace_all(country, pattern = "(?i)^u+n+i+[et]+[des]+[sd]*\\s+s+[at]+[aest]+[et]+[es]+[asd]*", "USA"))

countries <- countries %>%   
  mutate(country = str_remove_all(country, pattern = "(?i)^\\s*b+e+s+t+\\s+o+n+e+\\s+[:punct:]+\\s+"))


countries <- countries %>% 
  mutate(country = str_replace_all(country, pattern = "(?i)Alaska|California|North Carolina|Pittsburgh", "USA"))




# first attempts; banished! ----
#countries <- countries %>% 
#  mutate(country = str_replace_all(country, pattern = "(?i)T*h*e*[A-Za-z]*u+n+i+t+e+d+\\s+s+t+[ae]+t+e+s+\\s*of\\s*[Aa]merica[A-Za-z]*|(?i)[a-z]*USA[:punct:]*|[Nn][:punct:]\\s*America|(?i)usa+|(?i)america|^[A-Z][a-z]{3}[:punct:]{4}[A-Za-z]{6}|(?i)[:punct:]*[:symbol:]*m+[eu]+r+[ricka]*|(?i)unite*s*d*\\s*s*t*a*s*t*e*s+|New\\s+[JY][:alpha:]+$|^S+u+b........................|^I+\\sP+r+e+t+e+n+d+...........................................|[Tt]he\\s+[Bb]est\\s+[Oo]ne\\s*[:punct:]\\s*USA|[Uu]+[\\s\\.]+[Ss]+[\\s\\.]*\\s*[Aa]*\\s*\\.*|t[ae]+te+[ads]*$|USAUSA$|Alaska|California|North Carolina|Pittsburgh|^U+S+A+\\s*[:punct:]*[A-Za-z0-9]*[:punct:]*|(?i)U+S+A+\\s+O+f+A+", "USA"))

#countries <- countries %>%  # break it up a bit for sanity's sake
#  mutate(country = str_replace_all(country, pattern = "^[:alpha:]+\\s+S+U+S+A+|^U+S+A+U+S+A+|U+s+|U+S+A+\\s+[Oo]+f+\\s+A+|U+S+A+a+|U+S+A+\\s*[:alpha:]+(?i)u+s+a+\\s*u*s*a*|\\s*U+S+A+\\s*U+S+A+\\s*U*S*A*\\s*|\\s*U+S+A+\\s*U+S+A+\\s*U*S*A*\\s*|^[A-Za-z0-9]*\\s*U+S+A+\\s*[\\,\\.]*[A-Za-z0-9]*[\\.\\,]*[A-Za-z0-9]*|^U+S+A+\\s*...\\s*$|USA\\s*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*|(?i)\\s*u+s+a+u+s+a+\\s*|(?i)u+s+a+\\stell\\sanymore\\.\\.","USA"))

#countries <- countries %>%  # break it up a bit more...
 # mutate(country = str_replace_all(country, pattern = "U+S+A+U+S+A+|(?i)USAI[A-Za-z\\s]{1,65}|U+S+A+[:punct:]+", "USA"))

# Other general cleans: run this LAST ----

countries <- countries %>%     # remove numerics and various randoms.
  mutate(country = str_remove_all(country, pattern = "\\d+[:punct:]*\\d*$|(?i)[Nn]+o+t+\\s[Tt]*h*e*\\s*U+S+A+\\s+O*r*\\s*C+a+n+a+d+a+$|[Oo]ne\\sof\\sthe\\sbest\\sones|[Aa]tlantis|[Nn]arnia|[Dd]enial|[Ee]arth|[Ff]ear [Aa]nd [Ll]oathing|[Gg]od's [Cc]ountry|I [Dd]on't [Kk]now [Aa]nymore|[Ii]nsanity [Ll]ately|[Nn]everland|[Nn]ot [Tt]he USA [Oo]r Canada|[Oo]ne [Oo]f [Tt]he [Bb]est [Oo]nes|[Ss]ee [Aa]bove|[Ss]omewhere|(?i)subscribe to dm4uz3 on youtube|The Yoo Ess of Aaayyyyyy|there isn't one for old men|this one|Trumpistan|unhinged states|^A$|[Ee][Uu][Aa]|[Ee]urope|The |Republic [Oo]f\\s|U+d+|^\\s+|^[:punct:]|^[:symbol:]"))

countries <- countries %>% 
  mutate(country = na_if(country, ""))      # change blank cells to NA

countries <- countries %>% 
  drop_na(country)

#countries <- countries %>% 
#  mutate()

 # Capitalise country names;

countries <- countries %>% 
  mutate(country = str_to_title(country))


  

```

```{r}
candy_country <- countries %>% 
  
  filter(country == "Usa") %>%  
  
  group_by(name) %>% 
  
  summarise(top_usa = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_usa))

slice_max(candy_country, order_by = "value_numeric", n = 1)



candy_country <- countries %>% 
  
  filter(country == "Great Britain") %>%  
  
  group_by(name) %>% 
  
  summarise(top_uk = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_uk))

slice_max(candy_country, order_by = "value_numeric", n = 1)

# this was a draw - how to separate these?



candy_country <- countries %>% 
  
  filter(country == "Canada") %>%  
  
  group_by(name) %>% 
  
  summarise(top_canada = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_canada))

slice_max(candy_country, order_by = "value_numeric", n = 1)



candy_country <- countries %>% 
  
  filter(country != "Usa" & country != "Canada" & country != "Great Britain") %>% 
  
  group_by(name) %>% 
  
  summarise(top_other = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_other))

slice_max(candy_country, order_by = "value_numeric", n = 1)
  
```
The regex cleaning is not perfect, and even after several attempts still had some small inaccuracies.  However, 
the number of imperfectly tidied country names is a very small fraction of the overall and should not influence the
results by a great deal.


#### 8. So our result by country is as follows;

**USA:**    reese's peanut butter cups	        1973 votes

**UK:**     lindt truffle / rolos / toblerone     33 votes

**Canada:** kit kat	                             233 votes

**Other:**  kit kat	                              56 votes	


#### Note: a 3-way tie in the UK between lindt truffles, rolos and toblerones.
