---
title: "Halloween Sweets Analysis"
author: "Colin Scotland"
date: "29/03/2022"
output: html_document
---

The cleaning script for this analysis is `halloween_candy_cleaning_script.R`.  This
cleaning script performs the following processes;
- reads in three .csv files containing "trick or treat" data for 2015, 2016 and 2017
- removes unnecessary columns that are not needed for any analysis
- corrects the time format for 2017 and reduces all three data sets to show only the year
- changes the order of certain columns to set up for joining
- renames several columns to make the join easier
- performs a `full_join` to bring all the data into one data set while keeping all data
- cleans up the environment

The first process done upon exiting the cleaning script and setting up the libraries in this
Rmd file is to pivot the data into a longer form to allow easier analysis.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Set up libraries;
```{r}
library(tidyverse)
library(here)
library(assertr)
```

Set 'home' directory;
```{r}
here::here()
```

Source cleaning script;

```{r}
source("data_cleaning_scripts/halloween_candy_cleaning_script.R")
```

## 1. What is the total number of candy ratings given across three years.

### (Number of candy ratings, not the number of raters.  Don't count missing values).

```{r}
# Add an ID column to keep track after pivoting;
candy_all <- candy_all %>% 
  mutate(id = 1:nrow(candy_all), .before = "year")


candy_long <- candy_all %>% 
  pivot_longer(cols = c(-id, -year, -age, -gender, -country, -going_out))







candy_long_ratings <- candy_long %>% 
  
  mutate(num_ratings = if_else(value == "(?i)JOY" |
                               value == "(?i)MEH" |
                               value == "(?i)DESPAIR", as.integer(1), as.integer(0)), .after = "value") %>% 
  
  summarise(total_ratings = sum(!is.na(num_ratings)))

candy_long_ratings
  
```

#### There are **603883** ratings of sweets across the three years.  

These can also be broken down further to show how many of each type there were;

```{r}
candy_long %>% 
  
  select(value) %>% 
  
  distinct(value)
  
```

So there are only "JOY", "DESPAIR", "MEH" and NAs to deal with.

```{r}
candy_long_q1 <- candy_long %>% 
  
 # select(name, value) %>% 
  
  mutate(joy = (if_else(value == "JOY", 1, 0)), 
         despair = (if_else(value == "DESPAIR", 1, 0)), 
         meh = (if_else(value == "MEH", 1, 0)), 
         .after = "value") 

candy_long_q1 %>% 
  
  summarise(sum(joy, na.rm = TRUE), sum(despair, na.rm = TRUE), sum(meh, na.rm = TRUE))


```

Double check the count: 267672 + 260162 + 76049 = **603883**, which matches the original answer.


## 2.  What was the average age of people who are going out trick or treating?

As a start it would be good to have the age column formatted as integers;
```{r}
candy_all <- candy_all %>% 
  
  mutate(age = round(as.integer(age), digits = 1)) 
  

```
Note: the "junk" age entries have been coerced into NAs during the `as.integer` change, leaving
only numbers and NAs.  This is *a good thing* in this case.

In theory, a parent could take a baby trick or treating, and have answered the survey 
from their baby's perspective, so ages > 0 should be accepted.  At the other end of the 
spectrum, the oldest person in the world is 119 years old, so anyone aged 120 or over is lying. 

```{r}
candy_all <- candy_all %>% 
  filter(age > 0 & age < 120)

# and check there are no surprises;


candy_all %>% 
  filter(!is.na(age)) %>% 
  verify(age > 0 & age < 120) 
```

Now, discounting NAs we want to know what the average age is of people going out.

```{r}
candy_all %>% 
  
# what are the responses - are they only 'yes'/'no', or are there some oddities or colloquialisms in there?
  
  distinct(going_out)

# There are only 'Yes', 'No' and 'NA'.

candy_all %>% 
  
  filter(going_out == "Yes") %>% 
  summarise(avg_age_out = mean(age, na.rm = TRUE))
```

#### So the average age of people going out trick or treating is just over **35 years**.

## 3. What was the average age of people **not** going trick or treating?

Similarly to the last question;

```{r}
candy_all %>% 
  
  filter(going_out == "No") %>% 
  
  summarise(avg_age_out = mean(age, na.rm = TRUE))
```

#### The average age of people going out trick or treating is just over **39 years**.

## 4. For each joy, despair and meh, which candy bar received the most of these ratings?

*To be honest I think there's probably a more elegant way to present the results to this question *
*rather than having three separate tables arranged in descending order - if I get time I'll come back *
*to it. `slice_max(all_the_whatever, order_by = "xxx", n = 1)` wasn't working for me for some reason.*

```{r}

candy_rating_rates <- candy_long_q1 %>% 
  
  group_by(name, joy, meh, despair) %>% 
  
  summarise(all_the_joy = sum(as.integer(joy, na.rm = TRUE)), 
            all_the_meh = sum(as.integer(meh, na.rm = TRUE)), 
            all_the_despair = sum(as.integer(despair, na.rm = TRUE))
            )

candy_rating_rates %>% 
  
  select(name, all_the_joy) %>% 
  
  arrange(desc(all_the_joy))
  
candy_rating_rates %>% 
  
  select(name, all_the_meh) %>% 
  
  arrange(desc(all_the_meh))

candy_rating_rates %>% 
  
  select(name, all_the_despair) %>% 
  
  arrange(desc(all_the_despair))
 
  
```

#### The candy bar with the most ratings of each kind were;

**JOY**       | Any reese_s_peanut_butter_cups    | 7369 votes

**MEH**       | Lollipops                         | 1570 votes

**DESPAIR**   | Gum from baseball cards           | 7341 votes


## 5. How many people rated Starburst as despair?

```{r}
candy_rating_rates %>% 
  
select(name, all_the_despair) %>% 
  
filter(name == "starburst" & all_the_despair > 0)
```

#### There were **1990 votes of 'despair'** for Starburst.

## 6. What was the most popular candy bar by this rating system for each gender in the data set?

First, what entries do we have for gender;

```{r}
candy_long %>% 
  
  distinct(gender)
  
```
So we have to deal with 'NA', Male', 'Female', 'Other', and 'I'd rather not say'.

At least the formatting is consistent :)

```{r}
male_favourite <- candy_long_q1 %>% 
  
  select(gender, name, joy) %>% 
  
  filter(gender == "Male" & joy == 1)  %>% 
  
  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_male = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_male))

female_favourite <- candy_long_q1 %>% 
    
  select(gender, name, joy) %>% 
  
  filter(gender == "Female" & joy == 1) %>% 
  
  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_female = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_female))
  
other_favourite <- candy_long_q1 %>%

  select(gender, name, joy) %>% 
  
  filter(gender == "Other" & joy == 1) %>% 
  
  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_other = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_other))
  
nosay_favourite <- candy_long_q1 %>% 
  
  select(gender, name, joy) %>% 
  
  filter(gender == "I'd rather not say" & joy == 1) %>% 
  
  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_nosay = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_nosay))
  
na_favourite <- candy_long_q1 %>% 
  
  select(gender, name, joy) %>% 
  
  filter(is.na(gender) & joy == 1) %>% 

  group_by(name) %>% 
  
  drop_na(joy) %>% 
  
  summarise(total_joy_na = sum(as.integer(joy))) %>% 
  
  arrange(desc(total_joy_na))
```

```{r}

slice_max(male_favourite, "total_joy_male", n = 1)

slice_max(female_favourite, order_by = "total_joy_female", n = 1)

slice_max(other_favourite, order_by = "total_joy_other", n = 1)

slice_max(nosay_favourite, order_by = "total_joy_nosay", n = 1)

slice_max(na_favourite, order_by = "total_joy_na", n = 1)



```

#### So the most popular sweets by gender are;

**Male:**                  reese_s_peanut_butter_cups	            1545 votes

**Female:   **             reese_s_peanut_butter_cups	            837 votes

**Other: **                kit_kat	                              34 votes

**Prefer not to say:**     reese_s_peanut_butter_cups	            69 votes

**NA: **                   reese_s_peanut_butter_cups	            4890 votes	    



### *Now that I've got some results I've just realised that I haven't used the system *
### *defined by the task of setting despair = -1, meh = 0 and joy = +1.  So redo the *
### *question properly;*

```{r}
candy_long_q6 <- candy_long %>% 
  
  # set up the scores for ratings;
  
  mutate(value_numeric = case_when(value == "JOY" ~ as.numeric(1),
                                   value == "MEH" ~ as.numeric(0),
                                   value == "DESPAIR" ~ as.numeric(-1)))

```

The most popular sweets will be the ones with the highest scores after adding up all
the joy (+1) and despair (-1).


```{r}
q6_male <- candy_long_q6 %>% 
  
  filter(gender == "Male")  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_male = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(total_joy_male))




q6_female <- candy_long_q6 %>% 
  
  filter(gender == "Female")  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_female = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(total_joy_female))




q6_other <- candy_long_q6 %>% 
  
  filter(gender == "Other")  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_other = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(total_joy_other))




q6_nosay <- candy_long_q6 %>% 
  
  filter(gender == "I'd rather not say")  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_nosay = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(total_joy_nosay)) 




q6_na <- candy_long_q6 %>% 
  
  filter(is.na(gender))  %>% 
  
  group_by(name) %>% 
  
  summarise(total_joy_na = sum(value_numeric, na.rm = TRUE))  %>% 
  
  arrange(desc(total_joy_na))
```

That's the results grouped, so now to find the top scorers;

```{r}
slice_max(male_favourite, order_by = "q6_male", n = 1)

slice_max(female_favourite, order_by = "q6_female", n = 1)

slice_max(other_favourite, order_by = "q6_other", n = 1)

slice_max(nosay_favourite, order_by = "q6_nosay", n = 1)

slice_max(na_favourite, order_by = "q6_na", n = 1)
```

#### So the most popular sweets by gender using the *proper* system are;

**Male:**                  reese_s_peanut_butter_cups	            1545 votes

**Female:   **             reese_s_peanut_butter_cups	            837 votes

**Other: **                kit_kat	                              34 votes

**Prefer not to say:**     reese_s_peanut_butter_cups	            69 votes

**NA: **                   reese_s_peanut_butter_cups	            4890 votes	 

...which is exactly the same as the result from *not* using the defined "system", 
so at least they're consistent!

## 7. What was the most popular candy bar in each year?

```{r}

candy_long_2015 <- candy_long_q6 %>% 
  
  filter(year == "2015") %>%  
  
  group_by(name) %>% 
  
  summarise(top_2015 = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_2015))

slice_max(candy_long_2015, order_by = "value_numeric", n = 1)
  
  
```

```{r}

candy_long_2016 <- candy_long_q6 %>% 
  
  filter(year == "2016") %>%  
  
  group_by(name) %>% 
  
  summarise(top_2016 = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_2016))

slice_max(candy_long_2016, order_by = "value_numeric", n = 1)

```

```{r}

candy_long_2017 <- candy_long_q6 %>% 
  
  filter(year == "2017") %>%  
  
  group_by(name) %>% 
  
  summarise(top_2017 = sum(value_numeric, na.rm = TRUE)) %>% 
  
  arrange(desc(top_2017))

slice_max(candy_long_2017, order_by = "value_numeric", n = 1)
```

#### So the most popular sweets by year are;

**2015:**   reese_s_peanut_butter_cups	   4375 votes

**2016:**   kit_kat              	          920 votes

**2017:**   reese_s_peanut_butter_cups     1403 votes


## 8. What was the most popular candy bar by this rating for people in US, Canada,
UK and all other countries?

```{r}
countries <- candy_long %>% 
  
  distinct(country) %>% 
  arrange(country)

# Capitalise country names;
regex_1 <- countries %>% 
  mutate(country = str_to_title(country))
  


# UK cleanup ----

regex_1 <- regex_1 %>% 
  mutate(fixed = str_replace_all(country, pattern = "[Ee]n[dg]land|Scotland|(?i)u[:punct:]*k[:punct:]*|(?i)united king*dom", "Great Britain")) 

# Canada cleanup ----
regex_1 <- regex_1 %>% 
  mutate(fixed = str_replace_all(fixed, pattern = "(?i)\\s*cana*d*a*e*[:symbol:]*|(?i)cascadia\\s*\\s*|(?i)Soviet Canuckistan", "Canada"))


# USA cleanup ----

regex_1 <- regex_1 %>% 
  mutate(fixed = str_replace_all(fixed, pattern = "(?i)T*h*e*[A-Za-z]*u+n+i+t+e+d+\\s+s+t+[ae]+t+e+s+\\s*of\\s*[Aa]merica[A-Za-z]*|(?i)[a-z]*USA[:punct:]*|[Nn][:punct:]\\s*America|(?i)usa+|(?i)america|^[A-Z][a-z]{3}[:punct:]{4}[A-Za-z]{6}|(?i)[:punct:]*[:symbol:]*m+[eu]+r+[ricka]*|(?i)unite*s*d*\\s*s*t*a*s*t*e*s+|New\\s+[JY][:alpha:]+$|^S+u+b........................|^I+\\sP+r+e+t+e+n+d+...........................................|[Tt]he\\s+[Bb]est\\s+[Oo]ne\\s*[:punct:]\\s*USA|[Uu]+[\\s\\.]+[Ss]+[\\s\\.]*\\s*[Aa]*\\s*\\.*|t[ae]+te+[ads]*$|USAUSA$|Alaska|California|North Carolina|Pittsburgh|^U+S+A+\\s*[:punct:]*[A-Za-z0-9]*[:punct:]*|(?i)U+S+A+\\s+O+f+A+", "USA"))

regex_1 <- regex_1 %>% # break it up a bit for sanity's sake
  mutate(fixed = str_replace_all(fixed, pattern = "^[:alpha:]+\\s+S+U+S+A+|^U+S+A+U+S+A+|U+s+|U+S+A+\\s+[Oo]+f+\\s+A+|U+S+A+a+|U+S+A+\\s*[:alpha:]+(?i)u+s+a+\\s*u*s*a*|\\s*U+S+A+\\s*U+S+A+\\s*U*S*A*\\s*|\\s*U+S+A+\\s*U+S+A+\\s*U*S*A*\\s*|^[A-Za-z0-9]*\\s*U+S+A+\\s*[\\,\\.]*[A-Za-z0-9]*[\\.\\,]*[A-Za-z0-9]*|^U+S+A+\\s*...\\s*$|USA\\s*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*[.\\(\\)\\,]*","USA"))



# Other general cleans: run this LAST ----

regex_1 <- regex_1 %>%      # remove numerics and various randoms.
  mutate(fixed = str_remove_all(fixed, pattern = "\\d+[:punct:]*\\d*$|A [Tt]ropical [Ii]sland [Ss]outh [Oo]f [Tt]he [Ee]quator|[Nn]+o+t+\\s[Tt]*h*e*\\s*U+S+A+\\s+O*r*\\s*C+a+n+a+d+a+$|[Oo]ne\\sof\\sthe\\sbest\\sones|[Aa]tlantis|[Nn]arnia|[Dd]enial|[Ee]arth|[Ff]ear [Aa]nd [Ll]oathing|[Gg]od's [Cc]ountry|I [Dd]on't [Kk]now [Aa]nymore|[Ii]nsanity [Ll]ately|[Nn]everland|[Nn]ot [Tt]he USA [Oo]r Canada|[Oo]ne [Oo]f [Tt]he [Bb]est [Oo]nes|[Ss]ee [Aa]bove|[Ss]omewhere|(?i)subscribe to dm4uz3 on youtube|The Yoo Ess of Aaayyyyyy|there isn't one for old men|this one|Trumpistan|unhinged states|^A$|[Ee][Uu][Aa]|[Ee]urope|The |Republic [Oo]f\\s|U+d+|^\\s+|^[:punct:]|^[:symbol:]"))

regex_1 <- regex_1 %>% 
  mutate(fixed = na_if(fixed, ""))      # change blank cells to NA








  
```


